A matrix-free approach for finite-strain hyperelastic problems using geometric multigrid

### 2020-01-29 doc/IJNME/reply_to_reviewers.md

----

## Referee: 1

> Minor editing is required in figures (axis titles).

We have re-examined axis titles in all figures. We did not find any typos, but increased the font size of some of them to make more readable.

@@

> A uniform convention should be adopted for author names in the list of references.

The multiple first names were abbreviated the same way. This is the only non-uniformity we could find for the authors' names.

@@

> 2. To non-familiar Heidelberg-persons, IWR might be unknown. Please explain IWR on page 12.

We have replaced the system name "IWR" by "Westmere", which is the name of the CPU architecture and consistent with the use of "Cascade Lake" for the other system.

> 3. Section 6: Out of curiosity: how many nonlinear Newton steps are required to solve the problem?

With our settings, it takes 4 Newton-Raphson iterations to converge within a load step for the 2D problem and 3 iterations for the 3D setup.

> 4. In the conclusions, I would specifically refer again to Alg. 1,2,3 and give some suggestions which Alg. 1 finally is preferred and why.
> On page 16, just before Section 6.2, there are some suggestions. Are these the final conclusions w.r.t. the proposed Alg.?

If the linearization of the chosen material model allows to efficiently implement the action of the material part of the fourth-order spatial tangent stiffness tensor on the second-order symmetric tensor, then Algorithm 2 is recommended.
For applications within general frameworks with different geometrically non-linear constitutive models, Algorithm 3 provides the necessary flexibility while still being considerably faster than matrix-based strategies for polynomial degrees higher than one.

@@

Note that the higher arithmetic load for Algorithm 1 is not surprising as at each quadrature point we need to re-evaluate the Kirchhoff stress, and evaluate gradients with respect to both the referential and the current configuration.
Thus no advantage of the scalar caching can be deduced despite the higher arithmetic performance.

We have augmented the Conclusion section to better convey this message.

### 2020-01-29 doc/large_strain.tex

In this contribution, we proposed and numerically investigated three different matrix-free implementations of tangent operators for finite-strain elasticity with heterogeneous materials.
To the best of our knowledge, this is the first work that evaluates matrix-free sum factorization techniques on the partial differential equations arising in this case.
Among the three examined algorithms,
the implementation that caches the second-order Kirchhoff stress tensor and evaluates terms in the current configuration ({\color{red}Algorithm \ref{alg:mf_tensor2}}) is faster than
caching only a scalar at quadrature points ({\color{red}Algorithm \ref{alg:mf_scalar}}) or caching the material part of the fourth-order spatial tangent stiffness tensor together with the second-order Kirchhoff stress ({\color{red}Algorithm \ref{alg:mf_tensor4}}).
{\color{red}
Algorithm \ref{alg:mf_tensor2} is recommended when linearization of the adopted material model allows to efficiently implement the action of the material part of the fourth-order spatial tangent stiffness tensor on a second-order symmetric tensor.
Algorithm \ref{alg:mf_tensor4} does
}
not utilize the specific form of the material part of the stiffness tensor.
{\color{red}
In this case
}
the matrix-free implementation of the tangent operator is somewhat slower but more flexible because it can be applied to any constitutive model which
operates with the material part of the fourth-order spatial tangent stiffness tensor at the quadrature points.
{\color{red}
This implementation is still considerably faster than matrix-based strategies for polynomial degrees higher than one.
}

>>>>

In this contribution, we proposed and numerically investigated three different matrix-free implementations of tangent operators for finite-strain elasticity with heterogeneous materials.
To the best of our knowledge, this is the first work that evaluates matrix-free sum factorization techniques on the partial differential equations arising in this case.
Among the three examined algorithms,
the implementation that caches the second-order Kirchhoff stress tensor and evaluates terms in the current configuration ({\color{red}Algorithm \ref{alg:mf_tensor2}}) is faster than
caching only a scalar at quadrature points ({\color{red}Algorithm \ref{alg:mf_scalar}}) or caching the material part of the fourth-order spatial tangent stiffness tensor together with the second-order Kirchhoff stress ({\color{red}Algorithm \ref{alg:mf_tensor4}}).
{\color{red}
Algorithm \ref{alg:mf_tensor2} is recommended when linearization of the adopted material model allows to efficiently implement the action of the material part of the fourth-order spatial tangent stiffness tensor on a second-order symmetric tensor by two scalar quantities.
Algorithm \ref{alg:mf_tensor4} does
}
not utilize the specific form of the material part of the stiffness tensor.
{\color{red}
In this case
}
the matrix-free implementation of the tangent operator is somewhat slower but more flexible because it can be applied to any constitutive model which
operates with the material part of the fourth-order spatial tangent stiffness tensor at the quadrature points.
{\color{red}
This implementation is still considerably faster than matrix-based strategies for polynomial degrees higher than one.
}
